Won’t you be my neighboR?

### 2019-05-16 msu-workshop-2019.Rmd
update

1. Why learn R?
1. Getting started with R Studio
1. Computer programming basics
1. Visualizing data
1. Processing data
1. Modeling data
1. Creating reproducible reports
1. Learning and doing more with R

@@

* It is capable of carrying out basic and complex statistical analyses
* It is able to work with data small (*n* = 30!) and large (*n* = 100,000) efficiently
* It is a programming language and so is quite flexible
* There is a great, inclusive community of users and developers (and teachers)
* It is cross-platform, open-source, and freely-available

@@

### To download R Studio:

@@

---
## Creating Projects

@@

To create a project, in R Studio, navigate to "File" and then "New Directory".

Then, click "New Project". Choose a directory name for the project that helps you to remember that this is a project that involves data science in education; it can be convenient if the name is typed in `lower-case-letters-separated-by-dashes`, like that. You can also choose the sub-directory. If you are just using this to learn and to test out creating a project, you may consider placing it in your downloads or another temporary directory so that you remember to remove it later.

Even if you do not create a Project, you can always check where your working directory (i.e., where your R is pointing) is by running `getwd()`. To change it manually, run `setwd(desired/file/path/here)`.

@@

## Packages

@@

### Running Functions from Packages

@@

```{r}
help(package = dplyr)
```

@@

@@

@@

class: inverse, center, middle
# Visualizing data

# Visualizing distributions

# Comparing many means

# Relationships between two variables

# Relationships between three variables

---
class: inverse, center, middle
# Processing data

### Track Two: Welcome to the Tidyverse

You can installing and load the complete Tidyverse with:

**For more information on tidy data, check out [Hadley Wickhams's Tidy Data paper](http://vita.had.co.nz/papers/tidy-data.html).**

## Loading Data from Various Sources

In this section, we'll load data.

You might be thinking that an Excel file is the first that we would load, but there happens to be a format which you can open and edit in Excel that is even easier to use between Excel and R as well as SPSS and other statistical software, like MPlus, and even other programming languages, like Python. That format is CSV, or a comma-separated-values file.

The CSV file is useful because you can open it with Excel and save Excel files as CSV files. Additionally, and as its name indicates, a CSV file is rows of a spreadsheet with the columns separated by commas, so you can view it in a text editor, like TextEdit for Macintosh, as well. Not surprisingly, Google Sheets easily converts CSV files into a Sheet, and also easily saves Sheets as CSV files.

For these reasons, we start with - and emphasize - reading CSV files.

This next chunk of code downloads the file to your working directory. Run this to download it so in the next step you can read it into R. As a note: There are ways to read the file directory (from the web) into R. Also, of course, you could do what the next (two) lines of code do manually: Feel free to open the file in your browser and to save it to your computer (you should be able to 'right' or 'control' click the page to save it as a text file with a CSV extension).

student_responses_file_name <-
  paste0(getwd(), "/data/student-responses-data.csv")

```{r}
student_responses_file_name <-
  paste0(getwd(), "/data/student-responses-data.csv")
```

```{r}
download.file(
  url = student_responses_url,
  destfile = student_responses_file_name)
```

Now, in R Studio, you should see the downloaded file in the Files tab. This should be the case if you created a project with R Studio; if not, it should be whatever your working directory is set to. If the file is there, great. If things are *not* working, consider downloading the file in the manual way and then move it into the directory that the R Project you created it.

### Loading a CSV File

You may have noticed the hash symbol after the code that says `library(tidyverse`)`. It reads `# so tidyverse packages can be used for analysis`. That is a comment and the code after it (but not before it) is not run (the code before it runs just like normal). Comments are useful for showing *why* a line of code does what it does.

```{r, eval = TRUE}
# readr::write_csv(pisaUSA15::stu_quest, here::here("data", "pisaUSA15", "stu_quest.csv"))
student_responses <-
  read_csv("./data/student-responses-data.csv")
```

Woah, that's a big data frame (with a lot of variables with confusing names, to boot)!

Great job loading a file and printing it! We are now well on our way to carrying out analysis of our data.

### Loading Excel files

We will now do the same with an Excel file. You might be thinking that you can open the file in Excel and then save it as a CSV. This is generally a good idea. At the same time, sometimes you may need to directly read a file from Excel. Note that, when possible, we recommend the use of CSV files. They work well across platforms and software (i.e., even if you need to load the file with some other software, such as Python).

The package for loading Excel files, `readxl`, is not a part of the tidyverse, so we will have to install it first (remember, we only need to do this once), and then load it using `library(readxl)`. Note that the command to install `readxl` is grayed-out below: The `#` symbol before `install.packages("readxl")` indicates that this line should be treated as a comment and not actually run, like the lines of code that are not grayed-out. It is here just as a reminder that the package needs to be installed if it is not already.

Once we have installed readxl, we have to load it (just like tidyverse):

Of course, were this run, you can replace `my_data` with a name you like. Generally, it's best to use short and easy-to-type names for data as you will be typing and using it a lot.

Note that one easy way to find the path to a file is to use the "Import Dataset" menu. It is in the Environment window of R Studio. Click on that menu bar option, select the option corresponding to the type of file you are trying to load (e.g., "From Excel"), and then click The "Browse" button beside the File/URL field. Once you click on the, R Studio will automatically generate the file path - and the code to read the file, too - for you. You can copy this code or click Import to load the data.

### Loading SAV files

The same factors that apply to reading Excel files apply to reading `SAV` files (from SPSS). NOte that you can also read CSV file directly into SPSS and so because of this and the benefits of using CSVs (they are simple files that work across platforms and software), we recommend using CSVs when possible. First, install the package `haven`, load it, and the use the function `read_sav()`:

### Google Sheets

Finally, it can sometimes be useful to load a file directly from Google Sheets, and this can be done using the Google Sheets package.

```{r, eval = FALSE}
install.packages("googlesheets")
```

```{r, eval = FALSE}
library(googlesheets)
```

When you run the command below, a link to authenticate with your Google account will open in your browser.

```{r, eval = FALSE}
my_sheets <- gs_ls()
```

You can then simply use the `gs_title()` function in conjunction with the `gs_read()` function:

```{r, eval = FALSE}
df <- gs_title('title')
df <- gs_read(df)
```

### Saving Files

### Conclusion

We will detail the functions used to read every file in a folder (or, to write files to a folder).

## Processing Data

Now that we have loaded `student_responses` into an object, we can process it. This section highlights some common data processing functions.

[will add more on creating new variables, filtering grouping and summarizing, and joining data sets]

## Communicating / sharing results

R Markdown is a highly convenient way to communicate and share results. Navigate to "New File" and then "R Markdown". [add]

Then, click "Knit to PDF", "Knit to HTML", or "Knit to Word".

## Other foundational notes

### Configuring R Studio

There are a number of changes you *can* (but do not need to) make to configure R Studio. If you navigate to the Preferences menu in R Studio, you'll see a number of options you can change, from the appearance of the application to which windows appear where.

![optional caption text](images/save-workspace-reminder.jpg)

While this may seem like a dramatic step - never saving your workspace - it is the foundation for doing reproducible work and research using R Studio (and R). It also represents one of the biggest shifts from using software like Excel or SPSS, where most steps are not documented in code. This involves a shift from thinking that your most permanent and important part of an analysis is your data to thinking of the most important part as being the code: with the code, you can keep your data in its original form, process it, and then save a processed file, through running code. This also means that when you have to make a change to this code, you can re-run the entire analysis easily.

>>>>

1. Why learn R?
1. Getting started with R Studio
1. Visualizing data
1. Processing data
1. Learning and doing more with R
1. Modeling data (time permitting)
1. Computer programming basics (time permitting)
1. Creating reproducible reports (time permitting)

@@

* It is capable of carrying out basic and complex statistical analyses
* It is able to work with data small (*n* = 30!) and large (*n* = 100,000+) efficiently
* It is a programming language and so is quite flexible
* There is a great, inclusive community of users and developers (and teachers)
* It is cross-platform, open-source, and freely-available

@@

# To download R Studio:

@@

---
# Creating Projects

@@

To create a project, in R Studio, navigate to "File" and then "New Directory". Then, click "New Project".

@@

# Packages

@@

# Running Functions from Packages

@@

```{r, eval = FALSE}
help(package = dplyr)
```

@@

# Vignettes

```{r, eval = FALSE}
vignette(package = "dplyr")
vignette("dplyr", package = "dplyr")
```

# Using a specific function

@@

# Configuring R Studio

![optional caption text](images/save-workspace-reminder.jpg)
---

class: inverse, center, middle
# Processing data

# The Tidyverse

You can installing and load the complete tidyverse with:

**For more information on tidy data, check out [Hadley Wickhams's Tidy Data paper](http://vita.had.co.nz/papers/tidy-data.html).**
---

# Loading Data from Various Sources

student_responses_file_name <-
  paste0(getwd(), "./data/student-responses-data.csv")

# Step 1

# Step 2

```{r}
student_responses_file_name <-
  paste0(getwd(), "./data/student-responses-data.csv")
```

# Step 3

```{r, eval = FALSE}
download.file(
  url = student_responses_url,
  destfile = student_responses_file_name)
```

# Loading a CSV File

# Loading the file

```{r, eval = TRUE}
student_responses <-
  read_csv("./data/student-responses-data.csv")
```

# Loading Excel files

# Loading SAV files

# Saving Files

---
class: inverse, center, middle
# Processing Data
---

# The pipe: %>%

# Saving the results

# Renaming

@@
